Concrete compressive strength prediction with machine learning

### 2020-07-13 HarvardX_Data_Science_Capstone/CopyOfcapstone.Rmd
Harvard submission

bibliography: references.bib

@@

```{r download-data, echo=F, include=F}
# Data download
url_base <- "https://archive.ics.uci.edu"
url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")
n_inicial_samples <- nrow(dat)
colnames(dat)
```

@@

## Machine learning models

The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 and @Kuhn2008.

### Pre-processing and data separation

As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.

```{r dummy-var, echo=F, message=F, warning=F}
# Dummy variables
dummies <- dummyVars( ~ mix_app, data = dat)
dummyDat <- data.frame(predict(dummies, newdata = dat))
dummyDat$id <- dat$id
dat <- dat %>%
  select(-c(mix_app)) %>%
  full_join(., dummyDat)
```

The samples were separated based on age. One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}). For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-preparated-samples}.

```{r preparation, echo=F}
# Data preparation
names(dat) <- gsub(x = names(dat), pattern = "/", replacement = ".")  
dat_3 <- dat %>%
  select(-c("day_7", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_3",~"mpa")
dat_7 <- dat %>%
  select(-c("day_3", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_7",~"mpa")
dat_14 <- dat %>%
  select(-c("day_3", "day_7", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_14",~"mpa")
dat_28 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_28",~"mpa")
dat_56 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_100")) %>%
  drop_na() %>%
  rename_at("day_56",~"mpa")
dat_100 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_56")) %>%
  drop_na() %>%
  rename_at("day_100",~"mpa")
```

```{r table-preparated-samples, echo=F}
# Table - First 18 columns of the 6 first samples of 28 days
colNames = c("ID", "Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.", "MPa", "Class",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "App Mix" = 2)
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "1:1:2", "1:2:2")
caption <- "First 18 columns of the 6 first samples of 28 days"
dat_table <- dat_28[1:18]
kable(
    head(dat_table[order(dat_table$id),]),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

For each of the 6 sets, the existence or not of variables with variance close to zero and their subsequent removal was verified (\ref{show-nzv}). Many of the 23 variables added referring to the approximate concrete mix were removed due to this fact. In addition to them, in the case of the 7-day set, the fly ash variable was also removed. Then it was verified that there are no variables with high correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}). After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25 columns respectively for ages in increasing sequence.

```{r nzv, echo=F}
# Removing near zero variance columns
nzv_3 <- nearZeroVar(dat_3)
nzv_7 <- nearZeroVar(dat_7)
nzv_14 <- nearZeroVar(dat_14)
nzv_28 <- nearZeroVar(dat_28)
nzv_56 <- nearZeroVar(dat_56)
nzv_100 <- nearZeroVar(dat_100)
dat_3 <- dat_3[,-nzv_3]
dat_7 <- dat_7[,-nzv_7]
dat_14 <- dat_14[,-nzv_14]
dat_28 <- dat_28[,-nzv_28]
dat_56 <- dat_56[,-nzv_56]
dat_100 <- dat_100[,-nzv_100]
```

```{r cors, echo=F}
# High correlation variables verification
descr_cor_3 <- cor(select(dat_3, -c(mpa, id, class)))
descr_cor_7 <- cor(select(dat_7, -c(mpa, id, class)))
descr_cor_14 <- cor(select(dat_14, -c(mpa, id, class)))
descr_cor_28 <- cor(select(dat_28, -c(mpa, id, class)))
descr_cor_56 <- cor(select(dat_56, -c(mpa, id, class)))
descr_cor_100 <- cor(select(dat_100, -c(mpa, id, class)))
high_cor_3 <- sum(abs(descr_cor_3[upper.tri(descr_cor_3)]) > .999)
high_cor_7 <- sum(abs(descr_cor_7[upper.tri(descr_cor_7)]) > .999)
high_cor_14 <- sum(abs(descr_cor_14[upper.tri(descr_cor_14)]) > .999)
high_cor_28 <- sum(abs(descr_cor_28[upper.tri(descr_cor_28)]) > .999)
high_cor_56 <- sum(abs(descr_cor_56[upper.tri(descr_cor_56)]) > .999)
high_cor_100 <- sum(abs(descr_cor_100[upper.tri(descr_cor_100)]) > .999)
```

The stage of centralization and normalization of the variables was carried out later, together with the application of the models, as it is simpler to do this with the *caret* package. If performed at this time, it would be necessary to manually undo these transformations in the predictions. The *caret* allows you to transform before training the models and already transforms the results back.

Each data set was separated into test and training sets, *20%* and *80%* respectively (\ref{show-split}). The figure \ref{fig:dist-split} shows the distribution of data between the sets in relation to the compressive strength for each model (\ref{show-dist-split}).

```{r split, echo=F, message=F, warning=F}
# Test and train data split
reg <- list(
  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$mpa),
  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$mpa),
  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$mpa),
  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$mpa),
  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$mpa),
  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$mpa)
)
reg_seed <- c(
  1111, # 3
  1, # 7
  22, # 14
  11111, # 28
  111, # 56
  11 # 100
  )
split_reg <- function(n){
  set.seed(reg_seed[[n]], sample.kind="Rounding")
  createDataPartition(reg[[n]]$y, p = .8, list = F)
}
trainIndex_reg <- lapply(list(1,2,3,4,5,6), split_reg)
gen_dat <- function(n){
  regIndex <- trainIndex_reg[[n]]
  list(train = reg[[n]][regIndex,], test = reg[[n]][-regIndex,])
}
dats_reg <- lapply(list(1,2,3,4,5,6), gen_dat)
names(dats_reg) <- c("d3", "d7", "d14", "d28", "d56", "d100")
```

```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
# Distribution of test and train data
cap <- "Distribution of test and train data"
d_lab <- " days"
ylabel <- "Density"
g1 <- "Train"
g2 <- "Test"
dens <- function(d, n){
  dens <- full_join(
      d$train %>%
        select(y) %>%
        mutate(Group = g1, day = str_sub(n, 2)),
      d$test %>%
        select(y) %>%
        mutate(Group = g2, day = str_sub(n, 2))
    )
  dens
}
densities <- imap(dats_reg, dens) %>%
  reduce(rbind) %>%
  mutate(day_f = factor(day, levels=c('3','7','14','28', '56', '100')))
labs <- paste(c("3","7","14","28","56","100"), d_lab, sep="")
levels(densities$day_f) <- labs
densities %>%
  ggplot(aes(y, fill = Group)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~day_f, ncol=3) +
  xlab("MPa") +
  ylab(ylabel) +
  theme_bw() +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

### Performance measures

The performance evaluation of the models was performed by the Root Mean Square Error (*RMSE*). The *RMSE* is the measure used in all the works mentioned in the introduction and will allow the comparison of the models in the discussion.

### Naive models

Before creating the real models, for comparison purposes, naive models were created. They simply predict that the compressive strength of the test set is the average compressive strength of the training set (\ref{show-naive-model-reg}). In other words, naive models are simply the best guess possible. The results can be checked in the table \ref{tab:table-naive-model-reg}.

```{r naive-model-reg, echo=F, message=F, warning=F}
# Naive model
res <- function(d, n){
  data.frame(
    day = str_sub(n, 2),
    mean_mpa = mean(d$train$y)   
    )
}
ing_model <- imap(dats_reg, res)
get_rmse <- function(d, n){
   data.frame(
    rmse_train = RMSE(d$train$y, ing_model[[n]]$mean_mpa),
    rmse_test =  RMSE(d$test$y, ing_model[[n]]$mean_mpa),
    day = str_sub(n, 2) 
    )
}
rmses <- imap(dats_reg, get_rmse) %>%
  reduce(rbind)
ing_model_df <- ing_model %>% reduce(rbind)
df_performance_reg <- full_join(ing_model_df, rmses)
```

```{r table-naive-model-reg, echo=F}
# Table - Naive models
colNames = c("Age", "Mean $MPa$ (train)", "RMSE (train)", "RMSE (test)")
caption <- "Naive models"
kable(
    df_performance_reg,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

### Choice of algorithms

The *caret* [@caret] package exposes more than 200 different algorithms for building *machine learning* models. The package documentation presents an initial code [@modelsClusters] as a suggestion to select a portfolio of the most distinct algorithms possible in relation to some pre-selected algorithm, but for agility and due to technical limitations, it was chosen to use an algorithm with the highest probability to achieve the best possible result. According to @Fernandez2014, who compared 179 algorithms in 121 different databases, the algorithm most likely to achieve the best possible results is the *Parallel Random Forest* (called *prRF * in *caret*).

### Regression models

As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:

1. All *features*;
2. All *features* without the *dummy vars*;
3. Only the original *features*;
4. Only new *features*;
5. Only new *features*, without *dummy vars*;

Building a model for each of these configurations using the set at 28 days (\ref{show-test-models-reg}), showed that the best option is configuration 2, the *dummy vars* were completely discarded, but the other new variables were kept.
For illustrative purposes, the \ref{tab:table-done-samples-28} table shows the first 6 samples from the 28-day model training set. The samples of the other models, of the test and training sets are similar, the only difference being in the 7-day model, which excludes fly ash due to the variance close to zero, performed previously.

```{r test-models-reg, echo=F, warning=F, message=F, include=F}
# Features selection
data1 <- dats_reg$d28$train # full
data2 <- dats_reg$d28$train[c(1:13, 21)] # no dummy vars
data3 <- dats_reg$d28$train[c(1:7, 21)] # only original variables
data4 <- dats_reg$d28$train[c(8:21)] # only new variables
data5 <- dats_reg$d28$train[c(8:13, 21)] # only new varibels, no dummy vars
test1 <- dats_reg$d28$test # full
test2 <- dats_reg$d28$test[c(1:13, 21)] # no dummy vars
test3 <- dats_reg$d28$test[c(1:7, 21)] # only original variables
test4 <- dats_reg$d28$test[c(8:21)] # only new variables
test5 <- dats_reg$d28$test[c(8:13, 21)] # only new varibels, no dummy vars
# parRF
modelLookup("parRF")
control_parRF <- trainControl(method='repeatedcv',number=10,repeats=5,search='grid')
fit_parRF<- function(data, n) {
  set.seed(1, sample.kind = "Rounding")
  tunegrid_parRF<- expand.grid(mtry = seq(1, length(data), n))
  train(y ~ ., 
    data = data,
    preProcess = c("center","scale"),
    method='parRF', 
    tuneGrid=tunegrid_parRF,
    trControl=control_parRF)
}
fit_parRF1 <- fit_parRF(data1, 3)
ggplot(fit_parRF1)
min(fit_parRF1$results$RMSE) # 6.26108
p1_parRF1 <- predict(fit_parRF1, newdata = test1)
RMSE(p1_parRF1, test1$y) # 4.887264
fit_parRF2 <- fit_parRF(data2, 1)
ggplot(fit_parRF2)
min(fit_parRF2$results$RMSE) # 6.268789
p_parRF2 <- predict(fit_parRF2, newdata = test2)
RMSE(p_parRF2, test2$y) # 4.812994
fit_parRF3 <- fit_parRF(data3, 1)
ggplot(fit_parRF3)
min(fit_parRF3$results$RMSE) # 6.361905
p_parRF3 <- predict(fit_parRF3, newdata = test3)
RMSE(p_parRF3, test3$y) # 5.083467
fit_parRF4 <- fit_parRF(data4, 1)
ggplot(fit_parRF4)
min(fit_parRF4$results$RMSE) # 8.705984
p_parRF4 <- predict(fit_parRF4, newdata = test4)
RMSE(p_parRF4, test4$y) # 8.083962
fit_parRF5 <- fit_parRF(data5, 1)
ggplot(fit_parRF5)
min(fit_parRF5$results$RMSE) # 8.691755
p_parRF5 <- predict(fit_parRF5, newdata = test5)
RMSE(p_parRF5, test5$y) # 7.991537
```

```{r table-done-samples-28, echo=F}
# Table - First 6 samples of train data of the 28 days model
caption <- "First 6 samples of train data of the 28 days model"
colNames = c("Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "y")
dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","Ce.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "MPa")
colNames2 = c("Features"=13, "Outcome"=1)
kable(
    head(data2),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    add_header_above(header = colNames2, line = T, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

For each age set, a model was created using the *Parallel Random Forest* algorithm, previously defined (\ref{show-reg-models}). For each of the 6 models, the parameter *mtry* was optimized, and *repeated cross-validation * was performed, dividing into 10 or 30 parts and repeating 10 times.

```{r reg-models, echo=F, warning=F, message=F, include=F}
# Regression models
# no dummy vars:
data3 <- dats_reg$d3$train[c(1:13, 22)]
data7 <- dats_reg$d7$train[c(1:12, 19)]
data14 <- dats_reg$d14$train[c(1:13, 21)]
data28 <- dats_reg$d28$train[c(1:13, 21)]
data56 <- dats_reg$d56$train[c(1:13, 22)]
data100 <- dats_reg$d100$train[c(1:13, 23)]
test3 <- dats_reg$d3$test[c(1:13, 22)]
test7 <- dats_reg$d7$test[c(1:12, 19)]
test14 <- dats_reg$d14$test[c(1:13, 21)]
test28 <- dats_reg$d28$test[c(1:13, 21)]
test56 <- dats_reg$d56$test[c(1:13, 22)]
test100 <- dats_reg$d100$test[c(1:13, 23)]
get_trControl <- function(n, r){
  trainControl(method='repeatedcv',number=n,repeats=r,search='grid')
}
trControl3 <- get_trControl(30, 10)
trControl7 <- get_trControl(10, 10)
trControl14 <- get_trControl(30, 10)
trControl28 <- get_trControl(30, 10)
trControl56 <- get_trControl(30, 10)
trControl100 <- get_trControl(10, 10)
get_tuneGrid <- function(data){
  expand.grid(mtry = seq(1, length(data), 1))
}
tuneGrid3 <- get_tuneGrid(data3)
tuneGrid7 <- get_tuneGrid(data7)
tuneGrid14 <- get_tuneGrid(data14)
tuneGrid28 <- get_tuneGrid(data28)
tuneGrid56 <- get_tuneGrid(data56)
tuneGrid100 <- get_tuneGrid(data100)
set.seed(1, sample.kind = "Rounding")
fit_3 <- train(y ~ ., 
            data = data3,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid3,
            trControl=trControl3)
p_3 <- predict(fit_3, newdata = test3)
RMSE_test_3 <- RMSE(p_3, test3$y)
RMSE_test_3 # 3.31037
fit_3$bestTune # mtry = 6
set.seed(1, sample.kind = "Rounding")
fit_7 <- train(y ~ ., 
            data = data7,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid7,
            trControl=trControl7)
p_7 <- predict(fit_7, newdata = test7)
RMSE_test_7 <- RMSE(p_7, test7$y)
RMSE_test_7 # 4.361987
fit_7$bestTune # mtry = 2
set.seed(1, sample.kind = "Rounding")
fit_14 <- train(y ~ ., 
            data = data14,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid14,
            trControl=trControl14)
p_14 <- predict(fit_14, newdata = test14)
RMSE_test_14 <- RMSE(p_14, test14$y)
RMSE_test_14 # 4.620515
fit_14$bestTune # mtry = 13
set.seed(1, sample.kind = "Rounding")
fit_28 <- train(y ~ ., 
            data = data28,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid28,
            trControl=trControl28)
p_28 <- predict(fit_28, newdata = test28)
RMSE_test_28 <- RMSE(p_28, test28$y)
RMSE_test_28 # 4.716698
fit_28$bestTune # mtry = 11
set.seed(1, sample.kind = "Rounding")
fit_56 <- train(y ~ ., 
            data = data56,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid56,
            trControl=trControl56)
p_56 <- predict(fit_56, newdata = test56)
RMSE_test_56 <- RMSE(p_56, test56$y)
RMSE_test_56 # 5.939163
fit_56$bestTune # mtry = 8
set.seed(1, sample.kind = "Rounding")
fit_100 <- train(y ~ ., 
            data = data100,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid100,
            trControl=trControl100)
p_100 <- predict(fit_100, newdata = test100)
RMSE_test_100 <- RMSE(p_100, test100$y)
RMSE_test_100 # 5.851088
fit_100$bestTune # mtry = 8
```

# Results

The test *RMSE* for each model in ascending order of age was `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` and `r round(RMSE_test_100,2)` respectively. The table \ref{tab:table-reg-models} presents the details and results of each model (\ref{show-table-reg-models}). The figure \ref{fig:results-comparison} compares the actual and predicted values (\ref{show-results-comparison}), and the following tables show the best and worst results for each model (\ref{show-table-10}).

```{r table-reg-models, echo=F}
# Table - Models details
colNames <- c("Model", "mtry", "CV", "Repetitions", 
              "RMSE (train)", "RMSE (test)")
caption <- "Regression models results"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
dat_reg_models <- data.frame(
  dia = day,
  mtry = c(fit_3$bestTune$mtry, fit_7$bestTune$mtry, fit_14$bestTune$mtry,
           fit_28$bestTune$mtry, fit_56$bestTune$mtry, fit_100$bestTune$mtry),
  number = c(trControl3$number, trControl7$number, trControl14$number,
             trControl28$number,trControl56$number,trControl100$number),
  repeats = c(trControl3$repeats, trControl7$repeats, trControl14$repeats,
             trControl28$repeats,trControl56$repeats,trControl100$repeats),
  RMSE_train = c(min(fit_3$results$RMSE), min(fit_7$results$RMSE),
                  min(fit_14$results$RMSE), min(fit_28$results$RMSE),
                  min(fit_56$results$RMSE), min(fit_100$results$RMSE)),
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
# Figure - Models comparison
cap <- "Actual vs Predicted values in each model"
models <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
xlabel <- "Actual (MPa)"
ylabel <- "Predicted (MPa)"
preds <- list(p_3, p_7, p_14, p_28, p_56, p_100)
actuals <- list(test3$y, test7$y, test14$y, test28$y, test56$y, test100$y)
gen_res_df <- function(ind){
  data.frame(actual = actuals[[ind]], pred = preds[[ind]], model = models[[ind]])
}
res_df <- lapply(1:6, gen_res_df)
res_df <- bind_rows(res_df)
res_df$model <- factor(res_df$model,levels=models)
res_df %>%
  ggplot(aes(actual, pred)) +
  facet_wrap(~ model, ncol=3) +
  geom_point(alpha=0.5) +
  theme_bw() +
  geom_abline(slope=1, intercept=0) +
  xlab(xlabel) +
  ylab(ylabel)
```

```{r table-10, echo=F}
# Results tables of 10 best and worst results
colNames1 = c("Actual", "Predicted", "Error", "","Actual", "Predicted", "Error") 
colNames2 = c("Best 10"=3,"", "Worst 10"=3) 
caption_0 <- "Model of "
get_X <- function(pred, actual, X=10){
  diff <- abs(pred - actual)
  diff_2 <- pred - actual
  ind_min <- which(diff <= max(sort(diff, decreasing = F)[1:X]), arr.ind = T)
  ind_max <- which(diff >= min(sort(diff, decreasing = T)[1:X]), arr.ind = T)
  df_min <- data.frame(
    actual_min=actual[ind_min],
    pred_min=pred[ind_min],
    diff_min=diff[ind_min],
    diff_min_2=diff_2[ind_min]
  )
  df_max <- data.frame(
    actual_max=actual[ind_max],
    pred_max=pred[ind_max],
    diff_max=diff[ind_max],
    diff_max_2=diff_2[ind_max]
  )
  df_max <- df_max[order(-df_max$diff_max),]
  df_min <- df_min[order(df_min$diff_min),]
  df_null <- data.frame(null_col = rep(c(""), X))
  res <- cbind(df_min,df_null, df_max)
  res <- res %>% select(-c(diff_max, diff_min))
  rownames(res) <- c()
  res
}
gen_kable <- function(ind){
  df <- get_X(preds[[ind]], actuals[[ind]])
  caption <- paste0(caption_0,models[ind])
  kable(
    df,
    col.names = colNames1,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  column_spec(4, width = "1cm",) %>%
  add_header_above(header = colNames2, line = T, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position"))
}
gen_kable(1)
gen_kable(2)
gen_kable(3)
gen_kable(4)
gen_kable(5)
gen_kable(6)
```

# Discussion and conclusion

The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily. The alternative adopted to create a model for each set of age proved to be a valid method, managing to stratify to obtain specific results for each set. The studies cited in the introduction using the same dataset have similar results, as expected. The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .

```{r works-comparison, echo=F}
# Table - Comparison of other works
colNames = c("Author","Year" ,"Algorithm", "RMSE") 
caption <- "Comparison of other works"
works <- data.frame(
  autor = c("Pierobon", "Hameed", "Raj","Modukuru" ,"Alshamiri", "Abban"),
  ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
          "Gradient Boosting Regressor","Random Forest Regressor" ,
          "Regularized Extreme Learning Machine",
          "Support Vector Machines with Radial Basis Function Kernel"),
  RMSE = c("4.150", "4.736", "4.957","5.080","5.508", "6.105")
)
kable(
  works,
  col.names = colNames,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "l"
  )  %>%
    kable_styling(latex_options = c("HOLD_position"))
```

```{r results, echo=F}
# Table - Final results
colNames <- c("Model", "RMSE")
caption <- "Final result"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
dat_reg_models <- data.frame(
  dia = day,
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

Following the line of reasoning of this work, it can be performed with different algorithms, the results found here used only one (*Parallel Random Forest*), even though it was theoretically the "best" found, other algorithms can present even better results. Another option is to create an *ensemble* of various algorithms, just like @Pierobon2018, but with the separation of age sets proposed here. In addition, it can be performed with a larger dataset, ideally with the same number of samples in each age set, a more homogeneous distribution of compressive strength, and less variance between samples.

@@

\newpage

# Appendix 1 - Virtual environment {#appendix1}

## Operational system

```{r system, echo=FALSE}
v <- as.matrix(R.Version())
kable(v, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```

## Packages

```{r packages, echo=FALSE}
p <- matrix(
      c(
        "caret", as.character(packageVersion("caret")),
        "cowplot",as.character(packageVersion("cowplot")),
        "dplyr", as.character(packageVersion("dplyr")),
        "factoextra", as.character(packageVersion("factoextra")),
        "gdata", as.character(packageVersion("gdata")),
        "ggplot2", as.character(packageVersion("ggplot2")),
        "gridExtra",as.character(packageVersion("gridExtra")),
        "kableExtra",as.character(packageVersion("kableExtra")),
        "knitr",as.character(packageVersion("knitr")),
        "pastecs",as.character(packageVersion("pastecs")),
        "purrr", as.character(packageVersion("purrr")),
        "questionr", as.character(packageVersion("questionr")),
        "reshape2", as.character(packageVersion("reshape2")),
        "tidyr", as.character(packageVersion("tidyr")),
        "tidyverse", as.character(packageVersion("tidyverse"))
      ),
      ncol = 2,
      byrow=TRUE
    )
kable(p, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```
\newpage

# Appendix 2 - Code {#appendix2}

## Obtaining the data
### Data download

```{r show-download-data, ref.label='download-data', ref='show-download-data', eval=FALSE}
```

### Renaming the columns

```{r show-rename-dat-cols, ref.label='rename-dat-cols', ref='show-rename-dat-cols', eval=FALSE}
```

### Reordering the data

```{r show-reorder-dat, ref.label='reorder-dat', ref='show-reorder-dat', eval=FALSE}
```

### Defining column names and units

```{r show-col-names-and-units, ref.label='col-names-and-units', ref='show-col-names-and-units', eval=FALSE}
```

### Table - First samples

```{r show-first-samples, ref.label='first-samples', ref='show-first-samples', eval=FALSE}
```

## Data preparation
### Removing duplicated samples

```{r show-removing-duplicated-samples, ref.label='removing-duplicated-samples', ref='show-removing-duplicated-samples', eval=FALSE}
```

### Table - Samples with same composition

```{r show-similar-samples, ref.label='similar-samples', ref='show-similar-samples', eval=FALSE}
```

### Table - Same samples with different results

```{r show-similar-samples-2, ref.label='similar-samples-2', ref='show-similar-samples-2', eval=FALSE}
```

### Data cleaning

```{r show-initial-data-cleaning, ref.label='initial-data-cleaning', ref='show-initial-data-cleaning', eval=FALSE}
```

### Table - Previous samples after processing

```{r show-similar-samples-same-id, ref.label='similar-samples-same-id', ref='show-similar-samples-same-id', eval=FALSE}
```

### Figure - Compressive strength (MPa) vs age (days)

```{r show-boxplot, ref.label='boxplot', ref='show-boxplot', eval=FALSE}
```

### Figure - Principal component analysis - 90, 91 e 100 days

```{r show-pca-90-91-100, ref.label='pca-90-91-100', ref='show-pca-90-91-100', eval=FALSE}
```

### Figure - Compressive strength through time

```{r show-mpa-on-time, ref.label='mpa-on-time', ref='show-mpa-on-time', eval=FALSE}
```

### Joining 90, 91 and 100 days data

```{r show-join-90-91-100, ref.label='join-90-91-100', ref='show-join-90-91-100', eval=FALSE}
```

### Figure - Ages frequency

```{r show-freq-ages, ref.label='freq-ages', ref='show-freq-ages', eval=FALSE}
```

### Removing ages with frequency lower than 50

```{r show-remove-ages-lower-50, ref.label='remove-ages-lower-50', ref='show-remove-ages-lower-50', eval=FALSE}
```

```{r show-reorganizing-dat, ref.label='reorganizing-dat', ref='show-reorganizing-dat', eval=FALSE}
```

### Table - First 6 samples after reorganization

```{r show-new-features, ref.label='new-features', ref='show-new-features', eval=FALSE}
```

### Total samples

```{r show-total-samples-2, ref.label='total-samples-2', ref='show-total-samples-2', eval=FALSE}
```

### Adding New features

```{r show-new-features-2, ref.label='new-features-2', ref='show-new-features-2', eval=FALSE}
```

### Table - New features

```{r show-new-features-table, ref.label='new-features-table', ref='show-new-features-table', eval=FALSE}
```

## Data visualization
### Table - Descriptive statistics - continuous variables

```{r show-stat-summ, ref.label='stat-summ', ref='show-stat-summ', eval=FALSE}
```

### Figure - Descriptive statistics - categorical variables

```{r show-stat-summ-categorical, ref.label='stat-summ-categorical', ref='show-stat-summ-categorical', eval=FALSE}
```

### Figure - Correlation grouped by age

```{r show-correlation, ref.label='correlation', ref='show-correlation', eval=FALSE}
```

### Figure - Correlation over time

```{r show-correlation-mpa, ref.label='correlation-mpa', ref='show-correlation-mpa', eval=FALSE}
```

### Figure - Relationship between approximated mix, water, MPa and age

```{r show-mix-app-mpa, ref.label='mix-app-mpa', ref='show-mix-app-mpa', eval=FALSE}
```

### Figure - Relationship between concrete main features

```{r show-mix-mpa, ref.label='mix-mpa', ref='show-mix-mpa', eval=FALSE}
```

### Figure - Variables distribution

```{r show-vars-distribution, ref.label='vars-distribution', ref='show-vars-distribution', eval=FALSE}
```

### Figure - Variables distribution grouped by age

```{r show-vars-distribution-time, ref.label='vars-distribution-time', ref='show-vars-distribution-time', eval=FALSE}
```

### Figure - Principal component analysis on ingredients

```{r show-pca, ref.label='pca', ref='show-pca', eval=FALSE}
```

## Machine learning models
### Dummy variables

```{r show-dummy-var, ref.label='dummy-var', ref='show-dummy-var', eval=FALSE}
```

### Data preparation

```{r show-preparation, ref.label='preparation', ref='show-preparation', eval=FALSE}
```

### Table - First 18 columns of the 6 first samples of 28 days

```{r show-table-preparated-samples, ref.label='table-preparated-samples', ref='show-table-preparated-samples', eval=FALSE}
```

### Removing near zero variance columns

```{r show-nzv, ref.label='nzv', ref='show-nzv', eval=FALSE}
```

### High correlation variables verification

```{r show-cors, ref.label='cors', ref='show-cors', eval=FALSE}
```

### Test and train data split

```{r show-split, ref.label='split', ref='show-split', eval=FALSE}
```

### Distribution of test and train data

```{r show-dist-split, ref.label='dist-split', ref='show-dist-split', eval=FALSE}
```

### Naive model

```{r show-naive-model-reg, ref.label='naive-model-reg', ref='show-naive-model-reg', eval=FALSE}
```

### Table - Naive models

```{r show-table-naive-model-reg, ref.label='table-naive-model-reg', ref='show-table-naive-model-reg', eval=FALSE}
```

### Features selection

>>>>

# bibliography: references.bib
references:
- id: Yeh1998
  author:
  - family: Yeh
    given: I-Cheng
  year: 1998
  month: 12
  pages: 1797-1808
  title: Modeling of Strength of High-Performance Concrete Using Artificial Neural Networks
  volume: 28
  journal: Cement and Concrete Research
  DOI: 10.1016/S0008-8846(98)00165-3
- id: Hameed2020
  author:
  - family: Hameed
    given: Mohammed
  - family: Khalid
    given: Mohamed
   year: 2020
   month: 01
   pages: 323-335
   title: "Prediction of Compressive Strength of High-Performance Concrete: Hybrid Artificial Intelligence Technique"
   # isbn: 978-3-030-38751-8
   DOI: 10.1007/978-3-030-38752-5_26

@@

```{r download-data, echo=F, include=F}
# Data download
# url_base <- "https://archive.ics.uci.edu"
# url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
# download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")
n_inicial_samples <- nrow(dat)
colnames(dat)
```

@@

@@
